---
title: 'Project 2: Modeling, Testing, and Predicting'
author: "SDS348"
date: ''
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo = TRUE, eval = TRUE,fig.align="center",warning=FALSE,message=FALSE,fig.width=8, fig.height=5, linewidth=60)
options(tibble.width = 100,width = 100)
library(tidyverse)
```





##Introducing Dataset

```{r}
library(tidyverse)
library(dplyr)
library(plotROC)
library(pROC)
library(rstatix)
library(lmtest)
baseball <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/vcd/Baseball.csv")


data("baseball")

#altering dataset
baseball <- baseball %>%
mutate(league86 = ifelse(league86 == "N", 1, 0), div86 = ifelse(div86 == "E", 1, 0)) %>%
filter(complete.cases(.)) %>%
select(name1, name2, atbat86, hits86, homer86, runs86, rbi86, walks86, years, atbat, hits, homeruns, runs, rbi, walks, league86, div86, team86, sal87)

#number of observations
length(baseball$name1)

```

baseball is a data set with numerous variables regarding the yearly batting records for all major league baseball players. The variables include name 1 (player's first name), name 2 (player's last name), atbat86 (number of times at bat in '86), hits86 (number of hits in '86), homer86 (number of home runs in '86), runs86 (the number of runs scored by each player in '86), rbi86 (number of "runs batted in" in '86), walks86 (number of walks awarded to the hitter in '86), years (years played in the major league), atbat (number of times at bat throughout the career), hits (number of hits in career), homer (number of home runs in career), runs (the number of runs scored by each player in career), rbi (number of "runs batted in" in career), walks (number of walks awarded to the hitter in career), league86 (player's league in '86), div86 (player's division in '86), team86 (player's team in '86), sal87 (player's salar in '87 in thousands of dollars). There are 263 observations.


##MANOVA test
- **1. (15 pts)** Perform a MANOVA testing whether any of your numeric variables (or a subset of them, if including them all is unreasonable or doesn't make sense) show a mean difference across levels of one of your categorical variables (3). If they do, perform univariate ANOVAs to find response(s) showing a mean difference across groups (3), and perform post-hoc t tests to find which groups differ (3). Discuss the number of tests you have performed, calculate the probability of at least one type I error (if unadjusted), and adjust the significance level accordingly (bonferroni correction) before discussing significant differences (3). Briefly discuss MANOVA assumptions and whether or not they are likely to have been met (no need for anything too in-depth) (2).

```{r}
#MANOVA
manbase1<- (manova(cbind( homer86,hits86,runs86)~team86,data=baseball))
summary(manbase1)
summary.aov(manbase1)
```
This MANOVA test found a p.value of 0.1726, therefore it is insignificant. Therefore, I did not continue with running the ANOVA tests.



```{r}
group <- baseball$team86 
DVs <- baseball %>% select(homer86,hits86,runs86)
#Test multivariate normality for each group (null: assumption met)
sapply(split(DVs,group), mshapiro_test)

#If any p<.05, stop (assumption violated). If not, test homogeneity of covariance matrices

```
The pvalues that were less than 0.05 include Atl, Cal, Chi, Cin, Cle, Hou, LA, Mil, NY, Pit, SD, Sea, StL, and Tex. Assumptions were violated.


##Randomization Test

```{r}
#teams differ by salary
pairwise.t.test(baseball$sal87,baseball$team86, p.adj="none")
#teams differ by homers
pairwise.t.test(baseball$homer86,baseball$team86, p.adj="none")

baseball%>%select(team86,sal87,homer86)%>%pivot_longer(-1,names_to='DV', values_to='measure')%>%
  ggplot(aes(team86,measure,fill=team86))+geom_bar(stat="summary")+geom_errorbar(stat="summary", width=.5)+
  facet_wrap(~DV, nrow=2)+coord_flip()+ylab("")+theme(legend.position = "none")
``` 



##Linear Regression

```{r}
data("baseball")
sal87fix <- baseball$sal87-(mean(baseball$sal87, na.rm=T))
atbat86fix <- baseball$atbat86-(mean(baseball$atbat86, na.rm=T))
hits86fix <- baseball$hits86-(mean(baseball$hits86, na.rm=T))
homer86fix <- baseball$homer86-(mean(baseball$homer86, na.rm=T))
runs86fix <- baseball$runs86-(mean(baseball$runs86, na.rm=T))
rbi86fix <- baseball$rbi86-(mean(baseball$rbi86, na.rm=T))

#linear regression
baseballfit <- lm(sal87fix ~ atbat86fix + hits86fix + homer86fix + runs86fix + rbi86fix, data = baseball)
summary(baseballfit)

#graph 
baseballfit %>% ggplot(aes(sal87fix, 
    sal87fix, color = runs86fix)) + geom_point() + geom_smooth(method = "lm")


```

Holding all other variables constant, a player increasing their yearly RBI total by 1, increased their salary by approximately 6.74 thousand dollars, increasing their yearly at bat total by 1 decreased their salary by approximately 1.80 thousand dollars in 1987, increasing their hits total by 1 increased their salary by 6.30 thousand dollars in 1987, increasing their yearly homer total by 1 decreased their salary by approximately 3.43 thousand dollars in 1987, increasing their yearly run total by 1 increased their salary by approximately 2.03 thousand dollars in 1987 and for a player with 0 bats, hits, homeruns, runs, and rbis, the approximate salary for a player in 1987 was 3.49e-14 thousand dollars. Based on the adjusted R-squared value, 22.94% of the variation in the data can be explained. 

```{r}
#linearity, normality, and homoskedasticity 
#homoskedasticity test
library(sandwich)
library(lmtest)
bptest(baseballfit) #test assumption: violated!

coeftest(baseballfit, vcov = vcovHC(baseballfit))

#check assumptions
ggplot(mapping = aes(x=baseballfit$fitted.values, y=baseballfit$residuals)) + geom_point() + geom_hline(yintercept = 0) + geom_smooth()

ggplot(mapping = aes(sample = scale(baseballfit$residuals))) + geom_qq() + geom_qq_line()
```

running the "coeftest(baseballfit, vcov = vcovHC(baseballfit))" did not alter the t-test coefficients. The residuals are not linearly distributed as there is significant curving towards the beginning and end of the graph. The normality assumption is not met as the graph curves towards the end. The homoskedasticity assumption does not seem reasonable either.



##Rerun Regression

```{r}
### Bootstrapping residuals

fit<-lm(sal87fix ~ atbat86fix + hits86fix + homer86fix + runs86fix + rbi86fix, data = baseball) #fit model
resids<-fit$residuals #save residuals
fitted<-fit$fitted.values #save yhats/predictions

resid_resamp<-replicate(5000,{
new_resids<-sample(resids,replace=TRUE) #resample resids w/ replacement
baseball$sal87fix<-fitted+new_resids #add new resids to yhats to get new "data"
fit<-lm(sal87fix ~ atbat86fix + hits86fix + homer86fix + runs86fix + rbi86fix, data = baseball) 
coef(fit)
}) 

resid_resamp%>%t%>%as.data.frame%>%summarize_all(sd)


```
All of the variables were altered as compared to the previous SEs.

##Logistic Regression 
   
```{r}
library(plotROC)
baseballfitleag<-glm(league86~sal87+homer86, data=baseball, family="binomial")
coef(baseballfitleag)%>%exp%>%round(5)%>%data.frame

coeftest(baseballfitleag)

baseball$prob <- predict(baseballfitleag,type="response")

# predicted outcomes (if prob>.5, predict 0, otherwise predict 1)
baseball$predicted <- ifelse(baseball$prob>.5,"0","1")

```
From the coefficient estimates and their respective significance values, league and home runs in 1986 appear to be significant. 


```{r}
# confusion matrix
table(truth=baseball$league86, prediction=baseball$predicted)%>%addmargins

# accuracy: proportion of all cases that were correctly classified
(59+51)/263
#sensitivity: proportion of true that were correctly classified
51/131
# specificity: proportion of false cases that were correctly classified
59/132

# plot
library(plotROC)
baseball$logit<-predict(baseballfitleag)

```



```{r}
# ROC and AUC
baseballROC<- ggplot(baseball)+geom_roc(aes(d=league86, m=prob), n.cuts=0)
baseballROC

calc_auc(baseballROC)

baseball$logit<-predict(baseballfitleag)
baseball %>% ggplot(aes(logit, fill=league86)) + geom_density(alpha=.3)+geom_vline(xintercept=0, lty=2)
```
The AUC value is 0.613, which is not ideal. 


##More Logistic Regression

```{r}
library(glmnet)
set.seed(1234)
baseball_resp <- as.matrix(baseball$league86)  #response
baseball_preds <- model.matrix(league86 ~ ., data = baseball)[, -1]
head(baseball_preds)  #prediction

```


```{r}
cv <- cv.glmnet(baseball_preds, baseball_resp, family = "binomial")  #picks an optimal value for lambda through 10-fold CV

cv <- cv.glmnet(baseball_preds, baseball_resp, family = "binomial")
lasso_fit <- glmnet(baseball_preds, baseball_resp, family = "binomial", 
    lambda = cv$lambda.1se)
coef(lasso_fit)

predilasfit <- predict (lasso_fit, baseball_preds, tpye="response")
table(predict = as.numeric(predilasfit > 0.5), truth = baseball$league86)
```









